---
title: "CudaMon with Rcollectl"
output: rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{CudaMon with Rcollectl}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

This vignette shows how to combine GPU monitoring from `CudaMon` with CPU and
system metrics collected through `Rcollectl`.

At the moment the combined plotting helper lives in the package's experimental
area, so it must be loaded explicitly.

Install `Rcollectl` from Bioconductor before running the examples:

```{r install-rcollectl, eval = FALSE}
BiocManager::install("Rcollectl")
```

## Load packages and helper code

```{r load-rcollectl, eval = FALSE}
source(system.file("future", "viz.R", package = "CudaMon"))

if (!requireNamespace("Rcollectl", quietly = TRUE)) {
    stop(
        "Install Rcollectl with BiocManager::install('Rcollectl')",
        call. = FALSE
    )
}

library(CudaMon)
library(Rcollectl)
```

## Start CPU and GPU samplers together

`cl_start()` starts the `collectl`-based sampler, while `cm_start()` starts
the NVML sampler for GPU metrics.

```{r start-samplers, eval = FALSE}
cl_sampler <- cl_start()
cm_sampler <- cm_start()
```

## Run a workload and record event markers

The example below records three simple workflow steps while both samplers are
active.

```{r record-events, eval = FALSE}
Sys.sleep(2)
cm_timestamp(cm_sampler, "step 1")
Sys.sleep(2)
cm_timestamp(cm_sampler, "step 2")
Sys.sleep(2)
cm_timestamp(cm_sampler, "step 3")
Sys.sleep(2)
```

## Stop both samplers

Stop the GPU sampler first, then stop the `collectl` sampler and parse both
outputs.

```{r stop-samplers, eval = FALSE}
cm_stop(cm_sampler)
cl_stop(cl_sampler)

cm_session <- cm_parser(cm_sampler)
cl_session <- cl_parse(cl_result_path(cl_sampler))
```

## Plot CPU and GPU metrics together

`plot_usage()` combines:

- CPU and system metrics coming from `Rcollectl`
- GPU metrics coming from `CudaMon`
- event markers recorded with `cm_timestamp()`

The example below renders the combined plot in the Rome time zone.

```{r plot-combined-usage, eval = FALSE}
p <- plot_usage(cm_session, cl_session, tz = "Europe/Rome")
p
```

## Complete example


```{r complete-example, eval = FALSE}
source(system.file("future", "viz.R", package = "CudaMon"))

if (!requireNamespace("Rcollectl", quietly = TRUE)) {
    stop(
        "Install Rcollectl with BiocManager::install('Rcollectl')",
        call. = FALSE
    )
}

library(CudaMon)
library(Rcollectl)

cl_sampler <- cl_start()
cm_sampler <- cm_start()
Sys.sleep(2)
cm_timestamp(cm_sampler, "step 1")
Sys.sleep(2)
cm_timestamp(cm_sampler, "step 2")
Sys.sleep(2)
cm_timestamp(cm_sampler, "step 3")
Sys.sleep(2)
cm_stop(cm_sampler)
cl_stop(cl_sampler)

cm_session <- cm_parser(cm_sampler)
cl_session <- cl_parse(cl_result_path(cl_sampler))

pdf("session.pdf", width = 8, height = 6)
plot_usage(cm_session, cl_session, tz = "Europe/Rome")
dev.off()
```

## Session Info
```{r session-info, eval = TRUE}
sessionInfo()
```
