---
title: "Getting Started with CudaMon"
output: rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{Getting Started with CudaMon}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

`CudaMon` monitors NVIDIA GPU activity from R through NVML.
A typical session has four steps:

1. Start the sampler with `cm_start()`.
2. Run the workload you want to observe.
3. Add event markers with `cm_timestamp()` if you want annotated steps.
4. Stop the sampler, parse the output, and plot the session.

## Prerequisites

`CudaMon` requires:

- an NVIDIA GPU visible to the current R session
- a working NVML installation
- the package installed with its native code compiled successfully

If no GPU is available, the code below is still useful as a reference,
but it will not produce samples.

## Start a monitoring session

The simplest workflow starts a sampler tied to the current R session.
By default, `CudaMon` samples once per second and includes descendant
processes.

```{r load-cudamon, eval = TRUE}
library(CudaMon)

sampler <- cm_start()
```

You can customize the sampling interval or limit collection to a specific
device:

```{r start-custom-sampler, eval = TRUE}
sampler <- cm_start(
    period = 0.5,
    device_index = 0
)
```

## Annotate a workload

While the sampler is running, execute the code you want to profile.
Use `cm_timestamp()` to record named events that will later appear as
vertical markers in the plot.

```{r annotate-workload, eval = TRUE}
cm_timestamp(sampler, "initialization")

Sys.sleep(2)

cm_timestamp(sampler, "training")

Sys.sleep(2)

cm_timestamp(sampler, "evaluation")
```

## Stop and parse the session

When the workload is finished, stop the background sampler and parse the CSV
output into a `CudaMonSession`.

```{r stop-and-parse, eval = TRUE}
cm_stop(sampler)

session <- cm_parser(sampler)
session
```

The parsed session contains:

- `device_metrics`: device-level GPU samples over time
- `compute_processes`: active GPU processes observed during sampling
- `events`: event markers recorded with `cm_timestamp()`
- `paths`: the CSV and log files written by the sampler

## Visualize GPU activity

Use `cm_plot_usage()` for the built-in GPU plot.

```{r plot-gpu-usage, eval = TRUE}
if (length(session$compute_processes) > 0){
    p <- cm_plot_usage(session, tz = "Europe/Rome")
    p
}
```

This plot shows device utilization, GPU memory, temperature, and power over
time. If you recorded events, they are added as vertical dashed lines.


## Complete example

```{r complete-example, eval = TRUE}
library(CudaMon)

sampler <- cm_start()

Sys.sleep(2)
cm_timestamp(sampler, "step 1")
Sys.sleep(2)
cm_timestamp(sampler, "step 2")
Sys.sleep(2)
cm_timestamp(sampler, "step 3")
Sys.sleep(2)

cm_stop(sampler)

session <- cm_parser(sampler)
if (length(session$compute_processes) > 0){
    cm_plot_usage(session, tz = "Europe/Rome")
}
```

## Session Info
```{r session-info, eval = TRUE}
sessionInfo()
```
